Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-17 p2/axiom-def.tex
Improvements

% ax_separate_commute
\newcommand{\axaxseparatecommute}{
Commands on incomparable nodes commute: $\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$ where $n\unrel m$.
}

% ax_separate_nobreaks
\newcommand{\axaxseparatenobreaks}{
Commands on incomparable nodes also do not break all filesystems:
$\wrksx{\cxynv\cc\czwmv}$ where $n\unrel m$.
}

% ax_same_breaks
\newcommand{\axaxsamebreaks}{
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$ where $[y]\ne [z]$.
}

% ax_same_emptyseq
\newcommand{\axaxsameemptyseq}{
Commands on the same node are extended by an empty sequence if their types are compatible,
and their outer types represent an assertion command:
$\cxynv\cc\czwnv \eqext \emptyseq$ where $[y]=[z]$; and 
$[x]=[\empt]$ and $[w]=[\empt]$,
or $[x]=[\vald]$ and $[w]=[\vald]$.
}

% ax_same_singlec
\newcommand{\axaxsamesinglec}{
Command pairs on the same node are equivalent to a single command if their types are compatible,
and their outer types do not represent an assertion command:
$\cxynv\cc \czwnv \equiv \cxwnv$ where $[y]=[z]$; and
neither $[x]=[\empt]$ and $[w]=[\empt]$,
nor $[x]=[\vald]$ and $[w]=[\vald]$.
}

% ax_directchild_breaks
\newcommand{\axaxdirectchildbreaks}{
Commands on a parent and a child node break every filesystem
if the pair is not a construction pair
and the commands do not simply assert a directory at the parent
or an empty node at the child:
$\cxypnv\cc\czwnv \equiv \cbrk$ where $\cxypnv\cc\czwnv$ is not a construction pair, 
$\cxypnv\neq\cdda{\pn}$ and $\czwnv\neq\cbba{n}$.
}

% ax_directparent_breaks
\newcommand{\axaxdirectparentbreaks}{
Commands on a child and parent node break every filesystem
if the pair is not a destruction pair
and the commands do not simply assert an empty node at the child
or a directory at the parent:
$\cxynv\cc\czwpnv \equiv \cbrk$ where $\cxynv\cc\czwpnv$ is not a destruction pair,
$\cxynv\neq\cbba{n}$ and $\czwpnv\neq\cdda{\pn}$.
}

% ax_distantrel_breaks
\newcommand{\axaxdistantrelbreaks}{
Commands on distant relatives break all filesystems
if the commands do not simply assert a directory at the ancestor
or an empty node at the descendant:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
where $\nn\descendant n$ and $\nn\neq\parent(n)$, and $\cxynnv\neq\cdda{\nn}$ and $\czwnv\neq\cbba{n}$.
}

% ax_child_assert
\newcommand{\axaxchildassert}{
An assertion command can be freely added on a descendant node
next to a command that does not simply assert a directory:
\[ \cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n} \] 
where $\nn\descendant n$ and $\cxynnv\neq\cdda{\nn}$.
}

% ax_parent_assert
\newcommand{\axaxparentassert}{
An assertion command can be freely added on an ancestor node
next to a command that does not simply assert an empty node:
\[ \cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn} \]
where $\nn\descendant n$ and $\cxynv\neq\cbba{n}$.
}

>>>>

% ax_separate_commute
\newcommand{\axaxseparatecommute}{
Commands on incomparable nodes commute:
$\cxynv\cc\czwmv \equiv \czwmv\cc\cxynv$
when $n\unrel m$.
}

% ax_separate_nobreaks
\newcommand{\axaxseparatenobreaks}{
Commands on incomparable nodes also do not break all filesystems:
$\wrksx{\cxynv\cc\czwmv}$
when $n\unrel m$.
}

% ax_same_breaks
\newcommand{\axaxsamebreaks}{
Commands on the same node break every filesystem if their types are incompatible:
$\cxynv\cc\czwnv \equiv \cbrk$
when $[y]\ne [z]$.
}

% ax_same_emptyseq
\newcommand{\axaxsameemptyseq}{
Commands on the same node are extended by an empty sequence if their types are compatible,
and their outer types represent an assertion command:
$\cxynv\cc\czwnv \eqext \emptyseq$ 
when $[y]=[z]$, and
$[x]$ and $[w]$ are either both $[\empt]$ or both $[\vald]$.
}

% ax_same_singlec
\newcommand{\axaxsamesinglec}{
Command pairs on the same node are equivalent to a single command if their types are compatible,
and their outer types do not represent an assertion command:
$\cxynv\cc \czwnv \equiv \cxwnv$
when $[y]=[z]$, and
$[x]$ and $[w]$ are neither both $[\empt]$ nor both $[\vald]$.
}

% ax_directchild_breaks
\newcommand{\axaxdirectchildbreaks}{
Commands on a parent and a child node break every filesystem
if the pair is not a construction pair
and the commands do not simply assert a directory at the parent
or an empty node at the child:
$\cxypnv\cc\czwnv \equiv \cbrk$ when $\cxypnv\cc\czwnv$ is not a construction pair, 
$\cxypnv\neq\cdda{\pn}$ and $\czwnv\neq\cbba{n}$.
}

% ax_directparent_breaks
\newcommand{\axaxdirectparentbreaks}{
Commands on a child and parent node break every filesystem
if the pair is not a destruction pair
and the commands do not simply assert an empty node at the child
or a directory at the parent:
$\cxynv\cc\czwpnv \equiv \cbrk$ when $\cxynv\cc\czwpnv$ is not a destruction pair,
$\cxynv\neq\cbba{n}$ and $\czwpnv\neq\cdda{\pn}$.
}

% ax_distantrel_breaks
\newcommand{\axaxdistantrelbreaks}{
Commands on distant relatives break all filesystems
if the commands do not simply assert a directory at the ancestor
or an empty node at the descendant:
$\cxynnv\cc\czwnv \equiv \cbrk$
and $\czwnv\cc\cxynnv \equiv\cbrk$
when $\nn\descendant n$ and $\nn\neq\parent(n)$, and $\cxynnv\neq\cdda{\nn}$ and $\czwnv\neq\cbba{n}$.
}

% ax_child_assert
\newcommand{\axaxchildassert}{
An assertion command can be freely added on a descendant node
next to a command that does not simply assert a directory:
\[ \cbba{n}\cc\cxynnv \equiv \cxynnv \equiv \cxynnv\cc\cbba{n} \] 
when $\nn\descendant n$ and $\cxynnv\neq\cdda{\nn}$.
}

% ax_parent_assert
\newcommand{\axaxparentassert}{
An assertion command can be freely added on an ancestor node
next to a command that does not simply assert an empty node:
\[ \cdda{\nn}\cc\cxynv \equiv \cxynv \equiv \cxynv\cc\cdda{\nn} \]
when $\nn\descendant n$ and $\cxynv\neq\cbba{n}$.
}

### 2018-07-17 p2/axioms.tex
Improvements

We have included \cref{ax_directchild_breaks,ax_directparent_breaks}
for completeness, but we will use the more generic \cref{simple_distant_pairs},
which extends these rules to non-adjacent commands, although only in the case
of command sequences without any assertion commands.

@@

\begin{mydef}[$A\indep B$: Independent commands, sequences and sets of commands]\mlabel{def_indep}
Two commands $\alpha$ and $\beta$ 
are independent, and we write $\alpha\indep\beta$ if 
they commute and do not break all filesystems:
\[ \alpha\cc\beta \equiv \wrksx{\beta\cc\alpha}. \]
For two sequences or unordered sets of commands $A$ and $B$ we write $A\indep B$ if
for all $\alpha$ in $A$ and all $\beta$ in $B$, $\alpha\indep\beta$.
We also write $\alpha\indep B$ for $\{\alpha\}\indep B$.
\end{mydef}

It is intentional that we use the same symbol for independent commands
and sequences as for incomparable nodes. As
\namecref{incomparable_is_independent},
shows, these concepts are closely related.

@@

To prove this, we are going to prove the following lemma,
which characterizes independent commands in a more detailed way.

\begin{mylem}\mlabel{independent_details}
For two commands $\cxynv\indep\czwmv$ if and only if
one of the following holds:
\begin{itemize}
\item $n\unrel m$;
\item or $n=m$ and the commands are equivalent assertion or replacement commands;
\item or $n\descendant m$ or $m\descendant n$,
and either the command on the ancestor asserts a directory,
or the command on the descendant asserts an empty node (or both).
\end{itemize}
\end{mylem}

@@

\Cref{incomparable_is_independent} follows directly from \cref{independent_details}.
While equivalent commands can hardly be thought of as independent,
we retain this name for this relation as in most
cases we use it on pairs of different commands that are also not assertion commands.

>>>>

We have included \cref{ax_directchild_breaks,ax_directparent_breaks}
for completeness, but we will use the more generic \cref{simple_distant_pairs},
which extends these rules to non-adjacent commands, although only in the case
of command sequences without any assertion commands.
The \namecrefs{ax_separate_commute} can be derived from the filesystem model
relatively easily. Proofs are included in \cref{axiom_proof}.

@@

\begin{mydef}[$A\indep B$: Independent commands, sequences and sets of commands]\mlabel{def_indep}
Two commands $\alpha$ and $\beta$ 
are independent, written as $\alpha\indep\beta$, if 
they commute and do not break all filesystems:
\[ \alpha\cc\beta \equiv \wrksx{\beta\cc\alpha}. \]
For two sequences or unordered sets of commands $A$ and $B$ we write $A\indep B$ if
for all $\alpha$ in $A$ and all $\beta$ in $B$, $\alpha\indep\beta$.
We also write $\alpha\indep B$ for $\{\alpha\}\indep B$.
\end{mydef}

It is intentional that we use the same symbol for independent commands
as for incomparable nodes. As the
\namecref{incomparable_is_independent} below
shows, these concepts are closely related.

@@

\Cref{incomparable_is_independent} follows independently from
the following lemma,
which characterizes independent commands in a more detailed way.

\begin{mylem}\mlabel{independent_details}
For two commands $\cxynv\indep\czwmv$ if and only if
one of the following holds:
\begin{itemize}
\item $n\unrel m$,
\item $n=m$ and the commands are equivalent assertion or replacement commands,
\item either $n\descendant m$ or $m\descendant n$,
\emph{and} either the command on the ancestor asserts a directory,
or the command on the descendant asserts an empty node (or both).
\end{itemize}
\end{mylem}

@@

While equivalent commands can hardly be considered independent,
we retain this name for this relation as in most
cases we use it on pairs of different commands that are also not assertion commands.
